---
title: "Wine Club Reviews"
author: "Ken Wallace"
date: "May 27, 2016"
output: word_document
---

```{r}
setwd("C:/Users/conee/Dropbox/Data Science/WineClubReviews")
library(dplyr)
library(tidyr)
library(ggplot2)
library(chron)
```


# Read in data sets, merge data sets by club_id, and glimpse (dplyr) and view all data sets

```{r}
wineclubs <- read.csv("listings_export_051316.csv", stringsAsFactors = F)
glimpse(wineclubs)
View(wineclubs)
```
```{r}
wineclicks <- read.csv("tracked_clicks_full.csv", stringsAsFactors = F)
glimpse(wineclicks)
View(wineclicks)
```
```{r}
wine <- inner_join(wineclicks, wineclubs, by="club_id")
glimpse(wine)
View(wine)
```

# Assign a category name for each selcolor choice (from a lookup table provided separately)
# Future task: Try to merge the lookup table and convert this automatically
```{r}
wine$selcolor.name[wine$selcolor=="0"] <- "no choice"
wine$selcolor.name[wine$selcolor=="1"] <- "Red and White Mix only"
wine$selcolor.name[wine$selcolor=="2"] <- "Red-only option"
wine$selcolor.name[wine$selcolor=="3"] <- "White-only option"
wine$selcolor.name[wine$selcolor=="4"] <- "Red-only and White-only options"
wine$selcolor.name[wine$selcolor=="5"] <- "Red Wine"
wine$selcolor.name[wine$selcolor=="6"] <- "White Wine"
wine$selcolor.name[wine$selcolor=="7"] <- "Sparkling and Champagne"
wine$selcolor.name[wine$selcolor=="8"] <- "Mostly Reds"
wine$selcolor.name[wine$selcolor=="9"] <- "Mostly Reds option"
wine$selcolor.name[wine$selcolor=="10"] <- "Red and Sparkling"
wine$selcolor.name[wine$selcolor=="13"] <- "Rose"
wine$selcolor.name[wine$selcolor=="15"] <- "unknown"
wine$selcolor.name[wine$selcolor=="16"] <- "unknown"
```

# Select variables for new data frame
# Group data by club group and filter out clicks without sales (sales == total > 0)
# Goal: to pare down 143 variables to 19 variables to be used for balance of EDA

```{r}
wine.sub2 <- wine %>% 
  mutate(cdate = as.Date(converted_at, "%m/%d/%Y")) %>% 
  select(clubgroup_id.x, club_id, sale_amt, cdate, total, price, price_per_bottle, bottles, shipfreq, selcolor, shipprice, shipdesc, all_around, rating, selcolor.name, company_name, converted_at, keywords, session_id) %>% 
  group_by(clubgroup_id.x) %>% 
  filter(total>0) %>% 
  arrange(cdate)

head(wine.sub2, n=20)
```

```{r}
summary(wine.sub2)
```

# Identify top 20 clubgroups by revenue (descending)
# This includes 90% of revenue
```{r}
wine_top20 <- wine.sub2 %>% 
  group_by(clubgroup_id.x) %>% 
  summarise(revenue=sum(total)) %>% 
  arrange(desc(revenue)) %>% 
  slice(1:20)

wine_top20
```

# Bar chart of revenue by clubgroup_id
```{r}
ggplot(wine_top20, aes(as.factor(clubgroup_id.x), revenue)) + geom_bar(stat = "identity")

```

# Revenue by selcolor 
```{r}
wine_totalbyselcolor.name <- wine.sub2 %>% 
  group_by(selcolor.name) %>% 
  summarize(revenue=sum(total)) %>% 
  arrange(desc(revenue)) %>% 
  slice(1:20)

wine_totalbyselcolor.name
```

# Bar chart of revenue by selcolor
```{r}
ggplot(wine_totalbyselcolor.name, aes(selcolor, revenue)) + geom_bar(stat = "identity") + 
  scale_x_discrete(breaks=c(0:15))
```

# Revenue by bottles 
```{r}
wine_totalbybottles <- wine.sub2 %>% 
  group_by(bottles) %>%
  summarize(revenue=sum(total)) %>% 
  arrange(desc(revenue))

wine_totalbybottles
```

# Bar chart of revenue by bottles
```{r}
ggplot(wine_totalbybottles, aes(bottles, revenue)) + geom_bar(stat = "identity") + 
  scale_x_continuous(breaks=c(1:12))
```

# Revenue by average bottle price 
```{r}
wine_avgbottleprice <- wine.sub2 %>% 
  group_by(price_per_bottle) %>%
  summarize(revenue=sum(total)) %>% 
  arrange(desc(revenue))

wine_avgbottleprice
```

# Scatterplot of revenue by average bottle price
```{r}
ggplot(wine_avgbottleprice, aes(price_per_bottle, revenue)) + geom_point(alpha = 1) + 
  geom_smooth() +
  scale_x_continuous(breaks=seq(10,70,5)) + 
  scale_y_log10() +
  coord_cartesian(xlim = c(10,70))
```


# Plot of revenue by date
```{r}
ggplot(wine.sub3, aes(x = cdate, y = revenue, group = clubgroup_id.x, color = clubgroup_id.x)) + geom_line() + scale_x_date()
```


# Plot of cumulative total for club 27 by date
```{r}
wine.sub3.27 <- wine.sub3 %>% 
  filter(clubgroup_id.x=="27") %>% 
  na.omit()

p1 <- ggplot(wine.sub3.27, aes(x = cdate, y = cumsum(total)), group = year, color = year) + geom_line() + scale_x_date() + labs(title = "Club 27", x = "conversion date", y = "running total")

p1
```

# Plot of cumulative total for club 49 by date
```{r}
wine.sub3.49 <- wine.sub3 %>% 
  filter(clubgroup_id.x=="49") %>% 
  na.omit()

p2 <- ggplot(wine.sub3.49, aes(x = cdate, y = cumsum(total)), group = year, color = year) + geom_line() + scale_x_date() + labs(title = "Club 49", x = "conversion date", y = "running total")

p2
```

# Plot of cumulative total for club 171 by date
```{r}
wine.sub3.171 <- wine.sub3 %>% 
  filter(clubgroup_id.x=="171") %>% 
  na.omit()

p3 <- ggplot(wine.sub3.171, aes(x = cdate, y = cumsum(total)), group = year, color = year) + geom_line() + scale_x_date() + labs(title = "Club 171", x = "conversion date", y = "running total")

p3
```

# Plot of cumulative total for club 51 by date
```{r}
wine.sub3.51 <- wine.sub3 %>% 
  filter(clubgroup_id.x=="51") %>% 
  na.omit()

p4 <- ggplot(wine.sub3.51, aes(x = cdate, y = cumsum(total)), group = year, color = year) + geom_line() + scale_x_date() + labs(title = "Club 51", x = "conversion date", y = "running total")

p4
```

# Plot of cumulative total for club 70 by date
```{r}
wine.sub3.70 <- wine.sub3 %>% 
  filter(clubgroup_id.x=="70") %>% 
  na.omit()

p5 <- ggplot(wine.sub3.70, aes(x = cdate, y = cumsum(total)), group = year, color = year) + geom_line() + scale_x_date() + labs(title = "Club 70", x = "conversion date", y = "running total")

p5
```

# Plot of cumulative total for club 112 by date
```{r}
wine.sub3.112 <- wine.sub3 %>% 
  filter(clubgroup_id.x=="112") %>% 
  na.omit()

p6 <- ggplot(wine.sub3.112, aes(x = cdate, y = cumsum(total)), group = year, color = year) + geom_line() + scale_x_date() + labs(title = "Club 112", x = "conversion date", y = "running total")

p6
```

# Arrange plots on a grid (2 columns)
```{r}
grid.arrange(p1, p2, p3, p4, p5, p6, ncol=2)
```


##### OTHER EDA #####

# Create data frame of holiday period
# Future task: Evaluate sales within and outside of designated holiday periods

```{r}
wine.holiday <- data.table(start = c(as.Date(USThanksgivingDay(2010:2020))), 
                           end = as.Date(USNewYearsDay(2011:2021))-1)
wine.holiday
```

```{r}
wine.nonholiday #will be everything not in wine.holiday - still need to figure out how to do this.
```
