---
title: "Wine Club Reviews"
author: "Ken Wallace"
date: "May 27, 2016"
output: word_document
---

```{r}
library("dplyr")
library("tidyr")
library("ggplot2")
```


# Read in data sets, merge data sets by club_id, and glimpse (dplyr) and view all data sets

```{r}
wineclubs <- read.csv("listings_export_051316.csv")
#glimpse(wineclubs)
#View(wineclubs)
wineclicks <- read.csv("tracked_clicks_full.csv")
#glimpse(wineclicks)
#View(wineclicks)

wine <- merge(wineclicks, wineclubs, by="club_id")
#glimpse(wine)
#View(wine)

```

# select variables for new data frame
```{r}
wine_ <- wine %>% 
  select(clubgroup_id.x, club_id, name, category, sale_amt, total, converted_at, session_id, user_query, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords) %>% 

  group_by(clubgroup_id.x) %>% 
  mutate(revenue = sum(total)) %>% 
  
  select(clubgroup_id.x, club_id, name, category, sale_amt, total, revenue, converted_at, session_id, user_query, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords)

head(wine_)
```


```{r}
View(wine_)
```

### Total revenue in dataset


```{r}
summary(wine_$revenue)
summary(wine_$club_id)
```

# top 20 clubgroup_ids in revenue (descending) by clubgroup_id
```{r}
wine_top <- wine_ %>% 
  group_by(clubgroup_id.x) %>% 
  summarise(revenue=sum(total)) %>% 
  arrange(desc(revenue)) %>% 
  slice(1:20)

wine_top
```

# bar chart of number of sales by clubgroup_id
```{r}
ggplot(wine_top, aes(clubgroup_id.x, revenue)) + geom_bar(stat = "identity")

```

# revenue by selcolor 
```{r}
wine_totalbyselcolor <- wine_ %>% 
  group_by(selcolor) %>%
  summarize(revenue=sum(total)) %>% 
  arrange(desc(revenue))

wine_totalbyselcolor
```

# bar chart of revenue by selcolor
```{r}
ggplot(wine_totalbyselcolor, aes(selcolor, revenue)) + geom_bar(stat = "identity") + 
  scale_x_continuous(breaks=c(0:15)) +
  scale_y_log10(breaks=c(1,10,100,1000,10000,100000,1000000))
```

# revenue by bottles 
```{r}
wine_totalbybottles <- wine_ %>% 
  group_by(bottles) %>%
  summarize(revenue=sum(total)) %>% 
  arrange(desc(revenue))

wine_totalbybottles
```

# bar chart of revenue by bottles
```{r}
ggplot(wine_totalbybottles, aes(bottles, revenue)) + geom_bar(stat = "identity") + 
  scale_x_continuous(breaks=c(0:15))
```

# scatterplot of revenue by average bottle price
```{r}
ggplot(wine_, aes(price_per_bottle, revenue)) + geom_point() + 
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks=seq(10,70,5)) + 
  scale_y_log10() +
  coord_cartesian(xlim = c(10,70))
```

```{r}
ggplot(wine_, aes(price_per_bottle, revenue)) + geom_boxplot(aes(color="blue"))
```

# ***do not use for now*** total by club_id, grouped by clubgroup_name -- **do not use for now** 

```{r}
ggplot(wine_, aes(x = club_id, y = total)) + geom_point(alpha=.05) + facet_wrap(~clubgroup_name) +
  scale_y_log10()
```
Looks like the people linked from ggl and GoogleOrganic spend the most, then maybe Google UNK, BingOrganic, and then the Forbes article (not counting the blank/unknown sources)

# ***do not use for now*** price/bottle by club_id, grouped by clubgroup_name
```{r}
ggplot(wine_, aes(x = club_id, y = price_per_bottle)) + geom_point(alpha=.05) + facet_wrap(~clubgroup_name) + scale_y_log10()
```

# clicks by clubgroup_id
```{r}
wine_clubgroupid <- wine %>% 
  select(clubgroup_id.x, club_id, category, sale_amt, total, converted_at, session_id, user_query, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords) %>% 

  group_by(clubgroup_id.x) %>% 
  mutate(revenue = sum(total)) %>% 
  
  select(clubgroup_id.x, club_id, category, sale_amt, total, revenue, converted_at, session_id, user_query, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords)

head(wine_clubgroupid)
```

# sales by clubgroup_id - ***this df will be used to manipulate dates for time studies (next 2 chunks)***
```{r}
wine_clubgroupid.sales <- wine_ %>% 
  select(clubgroup_id.x, club_id, category, sale_amt, total, converted_at, session_id, user_query, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords) %>% 

  group_by(clubgroup_id.x) %>% 
  mutate(revenue = sum(total)) %>% 
  filter(total>0) %>% 
  summarise(revenue) %>% 
  
  select(clubgroup_id.x, club_id, category, sale_amt, total, revenue, converted_at, session_id, user_query, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords)

head(wine_clubgroupid.sales)
```

# first duplicate "converted_at" to cdate column, then convert to a date (this also drops the time stamp) and then separate out year/month
```{r}
wine_converted_dates <- wine_clubgroupid.sales %>% 
  mutate(cdate = as.Date(converted_at, "%m/%d/%Y")) %>% 
  select(clubgroup_id.x, club_id, category, sale_amt, total, revenue, converted_at, cdate, session_id, price, price_per_bottle, bottles, selcolor, billfreq, shipfreq, shipdesc, shipprice, duration, all_around, featured.x, clubgroup_name, list_name, company_name, keywords) %>% 
  separate(cdate, c("year", "month")) %>% 
  unite(YM, year:month, sep="-", remove=FALSE) %>% 
  arrange(YM) %>% 
  group_by(YM)

head(wine_converted_dates, n=20)
```

##### The following items are not resolved, and I would like to, but I'm stuck #####

# sales trends over time - i.e. by year, by month/year, and by holiday/non-holiday periods
# I figured out how get it to display year/month, by creating a new column, then arrange by that column (YM). Here is a bar chart, faceted by clubgroup_id. I'm on the right track.
```{r}
ggplot(wine_converted_dates, aes(YM, revenue)) + geom_bar(stat = "identity") + facet_wrap(~clubgroup_id.x, scales = "free_y")
```

# What I'd like are 2 line charts, where:

# 1. line chart - each clubgroup would be its own line, x would be month/year, and y would be *monthly* revenue, faceted by clubgroup
```{r}

```

# 2. line chart - each clubgroup would be its own line, x would be month/year, and y would be *cumulative* revenue, faceted by clubgroup
```{r}

```

# Then I'd like to figure out the following:
# 1. sales grouped by shipping cost - i.e. 3 buckets (free, under $x, over $x) - need to figure out what x is

# 2. quantity of sales at various price range buckets

# 3. For given session id's, including only ones containing a sale, which other clubs do customers look at before purchasing their chosen club?

# Lastly, it would be useful to figure out how to look at a single chart from a facet (i.e. facet wrap)

# After all of this, then I want to get into some of the predictive stuff.
